\LARGE \bf How Children Perceive and Interact with a Robot that Behaves Unexpectedly

### 2014-10-01 paper.tex
Implemented most of Pierre's comments

Using a playful domino game as the interaction scenario, we refer to this study
hereafter as the \textbf{``Domino Study''}.

@@

In the Domino Study, we analyse child-robot interaction with a robot that shows
unexpected behaviour. In a playful scenario which was set up in a laboratory
environment, 26 children aged 4-5 years (MD=4.46) were assembling a domino game together.
Each group consisted of two children and the \emph{Ranger} robot, which was used
to transport domino tiles between the two children.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Domino study}

@@

The interviews were analysed in a qualitative manner (we qualitatively
transcribed interviews, \ie we did not craft a full word-by-word transcript but
noted down any statement that was useful and relevant) and used to compute the
anthropomorphism index presented below.

@@

We have measured engagement by combining \emph{interaction}
data with the results of the semi-structured interviews.  Several of the
aforementioned coded actions can reflect engagement: \textit{explore, misuse,
gesture, touch, show,} and \textit{talk}. Also, \textit{look} at the
experimenter can be considered as engagement in the interaction: We assume that
children look at the experimenter when they are surprised and seek for help.
This behaviour reflects that they want to make sense of the robot's behaviour and
that they notice it as ``strange''. Further, we can take into account how they
refer to Ranger and describe the robot (as well as their experience) in the
interviews.  Specifically, we used the constructs \textit{ascribe intention,
ascribe cognitive abilities, ascribe emotional state}, and \textit{ascribe
moral standing} (Table~\ref{tab:domino-questions}) to assess how far they
engaged with the robot.

To account for the fact that anthropomorphism arises in an interaction, we tried
to bring both children's perception of the robot (post-measurement) and their
behaviour toward it (in-the-moment measurement) together. We did so by building
a qualitative anthropomorphism index~\cite{fink2014dynamics}. We quantify the
index by attributing points for each anthropomorphic \emph{perception} of the
robot (max.  13 points) and for specific kinds of human-like \emph{behaviour}
toward the robot (max.  3 points). The index included the following aspects:\\

@@

There are certainly limitations to this scoring scheme. For instance, the
balance between perception and behaviour aspects is questionable (13 to 3
points). Also, we did not consistently assign 1 point to each item, but assigned
points between 0.5 and 2 points. We did this because we found that different
items reflect a higher level of anthropomorphic perception of the robot than
others (for instance, ascribing the ability to see and hear was suggested by our
study setup, and we cannot be sure that it really reflects anthropomorphism).

@@

Overall, we can notice that not all children perceived the misbehaviour of the
robot as we had intended and we have to keep this in mind while interpreting the
data. Also \cite{leite_long-term_2013} found in her study that when children did
not understand an action of the robot, they tended to view it as a mistake,
rather than interpreting the robot's behaviour as a deliberative action.

@@

\paragraph{Engagement}
In terms of \textbf{engagement}, the novelty of the robot certainly plays a role
here. The findings here do not directly address the issues
of long-term usage but concern short-term \textit{engagement}, which is a
pre-requisite for long-term usage.

@@

After they had interacted with the misbehaving robot, we asked children again.
Now, still 11 children (previously 12) believed Ranger would always obey to
them. When analysing the answers of each child separately, there were 2 children
in the \textit{disobey} condition that changed their answer from \textit{yes} to
\textit{no}; however also 1 child in the \textit{lost} condition that switched
their answer in the opposite way (strangely). There was a more remarkable change
when asking children whether Ranger could do something silly. In the second
interview, 8 children (previously~3) answered Ranger could do something silly. 1
child in the \textit{disobey} condition had changed their answer, and each 2 in
the \textit{mistake} and \textit{lost} condition.

We can summarise that even with the \textbf{unexpectedly behaving robot}
children do \textbf{not necessarily ascribe intention} to the robot. The
unexpected robot behaviour impacted only some children's attributions of
intention to the robot.  It seemed that some children did not interpret the
robot's misbehaviour as intentional but more like a technical problem or
mistake. For instance, even after the robot misbehaved by \textit{disobeying},
the majority of the children in this condition was still convinced that the
robot could not do a silly thing.  It it interesting to note that children tend
to ascribe cognitive abilities to the robot, like the ability to see and hear
but not intention. We interpret that children may perceive the robot as being
able to process sensory information but that it is not able to make decisions on
its own.

@@

To obtain the anthropomorphism index for a pair, we first calculated the index
per child and then took the average of the two children in one group.  The group
indexes varied between 3.25 and 10.75 points with an average of 7.5 points
(SD~=~2.5), which is 47~\% of the maximum possible index of 16 points.
Intra-group variations were interesting: In 7 of the 13 groups, the
anthropomorphism indexes for both children were similar (difference less than
1.5 points), which means that both children anthropomorphized the robot to a
similar degree. This agreement among children happened for both higher and lower
indexes. In 6 of the 13 groups, the anthropomorphism indexes varied in more than
2 points, which means that one of the children anthropomorphized the robot more
than the other one.

@@

We hypothesised that children who interact a lot and are probably more engaged
with the robot also perceive the robot as more human-like. Data suggests the
opposite, however. As shown in Figure~\ref{fig:domino-anthropo-interaction}, we
found a significant \textit{negative} correlation between the count of
engagement actions (per group) and the qualitative anthropomorphism index
(r(11)=-0.56, p=.05).

%	(b=-.05, t(11)=-2.22, p=.048).

The overall model with the count of engagement actions (per group) predicts a
significant proportion of the qualitative anthropomorphism index (adjusted
R$^2$=.25, F(1,11)=4.9, p=.05). This means that the more a group showed
engagement in the interaction, the less they anthropomorphized the robot. This
is a key results, which was against our initial assumption. A possible
interpretation is that children who interact more with the robot understand better
how it works, they are more familiar with it, and as such the robot appears less
``mystical'' to them, so there is not a big need to anthropomorphize it. On the
contrary, the cluster of groups that do not interact much but anthropomorphize
the robot more, is quite homogeneous. The negative correlation between the
number of engagement actions and the anthropomorphism index suggests that
children who interact more with the robot tend to anthropomorphize it less. This
implies that anthropomorphism could fade out after some time (or rather after
some interaction). This evokes the question how far anthropomorphism (as a
special kind of social engagement) really helps in sustaining interaction. This
is a critical point because most of the short-term investigations suggest that
anthropomorphic design and human social cues emitted by a robot foster
engagement and acceptance. What if this is not true for continued interaction,
and thus for the long-term?  We have to be careful about our interpretation, due
to the small sample size with which we obtained these findings. We suggest to
investigate the aspect further in future research.

@@

    \caption[Correlation of Anthropomorphic Perception of the Robot and
    Interaction]{\small \textbf{Correlation of anthropomorphic perception of the
    robot and engagement actions per group.} A scatter plot of the count of engagement
    actions (without \emph{explore}) and the \textit{anthropomorphic perception} (score per
    group) shows a negative correlation. The score for the anthropomorphic
    perception (\textit{y-axis}) does not take into account the behavioural aspect of
    the anthropomorphism index, as this is part of the interaction
    (\textit{x-axis}).  There seem to be two clusters of groups: those who interact
    more with the robot and anthropomorphize it less, and those who interact less
    but anthropomorphize the robot more.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% double blind review
%\section*{Acknowledgements}
%
%This research was supported by the Swiss National Science Foundation through the
%National Centre of Competence in Research Robotics.

>>>>

@@

We analyse child-robot interaction with a robot that shows unexpected behaviour.
In a playful scenario which was set up in a laboratory environment, 26 children
aged 4-5 years (MD=4.46) were assembling a domino game together. Each group
consisted of two children and the \emph{Ranger} robot, which was used to
transport domino tiles between the two children.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Research Methodology}

@@

We did not craft a full word-by-word transcript but
noted down any statement that was useful and relevant) and used to compute the
anthropomorphism index presented below.

@@

We have measured engagement by combining \emph{interaction} data with the
results of the semi-structured interviews.  Several of the aforementioned coded
actions can reflect engagement: \textit{explore, misuse, gesture, touch, show,}
and \textit{talk}. Also, \textit{look} at the experimenter can be considered as
engagement in the interaction: We assume that children look at the experimenter
when they are surprised and seek for help.  This behaviour reflects that they
want to make sense of the robot's behaviour. Further, we can take into account
how they refer to Ranger and describe the robot (as well as their experience) in
the interviews.  Specifically, we used the constructs \textit{ascribe intention,
ascribe cognitive abilities, ascribe emotional state}, and \textit{ascribe moral
standing} (Table~\ref{tab:domino-questions}) to assess how far they engaged with
the robot.

To account for the fact that anthropomorphism arises in an interaction, we bring
both children's perception of the robot (post-measurement) and their behaviour
toward it (in-the-moment measurement) together. We did so by building a
anthropomorphism index~\cite{fink2014dynamics}. We quantify the
index by attributing points for each anthropomorphic \emph{perception} of the
robot (max.  13 points) and for specific kinds of human-like \emph{behaviour}
toward the robot (max.  3 points). The index included the following aspects:\\

@@

There are certainly limitations to this scoring scheme. For instance, the
balance between perception and behaviour aspects is questionable and the
weighting of each item can be discussed: we did not consistently assign 1 point
to each item, but assigned points between 0.5 and 2 points depending on our
perception of how a given item reflect on a higher level of anthropomorphic
perception of the robot (for instance, ascribing the ability to see and hear was
suggested by our study setup, and we cannot be sure that it really reflects
anthropomorphism). We believe however that the approach remains valid,
considering that the anthropomorphism index is primarily a \emph{comparison}
metric and not an absolute value.

@@

Not all children perceived the misbehaviour of the robot as we had intended and
we have to keep this in mind while interpreting the data. Also
\cite{leite_long-term_2013} found in her study that when children did not
understand an action of the robot, they tended to view it as a mistake, rather
than interpreting the robot's behaviour as a deliberative action.

@@

\paragraph{Engagement}
In terms of \textbf{engagement}, the novelty of the robot certainly plays a role
here. The findings do not directly address the issues
of long-term usage but concern short-term \textit{engagement}, which is a
pre-requisite for long-term usage.

@@

After they had interacted with the misbehaving robot, we asked children again.
Still 11 children believed Ranger would always obey to them. There was a
more remarkable change when asking children whether Ranger could do something
silly. In the second interview, 8 children (previously~3) answered Ranger could
do something silly. 1 child in the \textit{disobey} condition had changed their
answer, and each 2 in the \textit{mistake} and \textit{lost} condition.

We can summarise that even with the \textbf{unexpectedly behaving robot}
children do \textbf{not necessarily ascribe intention} to the robot. The
unexpected robot behaviour impacted only some children's attributions of
intention to the robot.  It seemed that some children did not interpret the
robot's misbehaviour as intentional but more like a technical problem or
mistake. For instance, even after the robot misbehaved by \textit{disobeying},
the majority of the children in this condition was still convinced that the
robot could not do a silly thing.  It is however interesting to note that children tend
to ascribe cognitive abilities to the robot, like the ability to see and hear
but not intention. We interpret that children may perceive the robot as being
able to process sensory information but that it is not able to make decisions on
its own.

@@

To obtain the anthropomorphism index for a pair, we first calculated the index
per child and then took the average of the two children in one group.  The group
indexes varied between 3.25 and 10.75 points with an average of 7.5 points
(SD~=~2.5), which is 47~\% of the maximum possible index of 16 points.
Intra-group variations were interesting: In 7 of the 13 groups, the
anthropomorphism indexes for both children were similar (difference less than
1.5 points). This agreement among children happened for both higher and lower
indexes. In 6 of the 13 groups, the anthropomorphism indexes varied in more than
2 points, which means that one of the children anthropomorphized the robot more
than the other one.

@@

We hypothesised that children who interact a lot and are probably more engaged
with the robot also perceive the robot as more human-like. Data suggests the
opposite, however. As shown in Figure~\ref{fig:domino-anthropo-interaction}, 
%(we even found a significant \textit{negative} correlation between the count of
%engagement actions (per group) and the qualitative anthropomorphism index --
%r(11)=-0.56, p=.05 -- but the sample size is too small to report it).  (b=-.05,
%t(11)=-2.22, p=.048).  The overall model with the count of engagement actions
%(per group) predicts a significant proportion of the qualitative
%anthropomorphism index (adjusted R$^2$=.25, F(1,11)=4.9, p=.05). This means
%that 
the more a group showed engagement in the interaction, the less they
anthropomorphized the robot. This is a key result, which was against our
initial assumption. A possible interpretation is that children who interact more
with the robot understand better how it works, they are more familiar with it,
and as such the robot appears less ``mystical'' to them, so there is not a big
need to anthropomorphize it. On the contrary, the cluster of groups that do not
interact much but anthropomorphize the robot more, is quite homogeneous, and may
reflect a certain fear of interacting with a robot that would look to them to
human-like.
%The negative correlation between the number of engagement actions and the
%anthropomorphism index suggests that children who interact more with the robot
%tend to anthropomorphize it less. This implies that anthropomorphism could fade
%out after some time (or rather after some interaction).
This raises the question how far anthropomorphism (as a special kind of social
engagement) really helps in sustaining interaction. This is a critical point
because most of the short-term investigations suggest that anthropomorphic
design and human social cues emitted by a robot foster engagement and
acceptance. What if this is not true for continued interaction, and thus for the
long-term? We need to remain modest here: while we found a significant
\emph{negative} correlation between engagement and anthropomorphism in the data
pictured on Figure~\ref{fig:domino-anthropo-interaction} (r(11)=-0.56, p=.05),
we have to be careful about our interpretation, due to the small sample size (13
pairs), and we suggest to investigate the aspect further in future research.

@@

    \caption{\small \textbf{Anthropomorphic perception of the robot versus
        engagement actions per pairs.} A scatter plot of the count of engagement
        actions (without \emph{explore}) and the \textit{anthropomorphic
        perception} (score per group) evidence one cluster of
        high-anthropomorphizers/low-interactors and a another group of pairs who
        interact more with the robot and anthropomorphize it less. The score for
        the anthropomorphic perception (\textit{y-axis}) does not take into
        account the behavioural aspect of the anthropomorphism index, as this is
        part of the interaction (\textit{x-axis}).}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
%
% double blind review
%This research was supported by the Swiss National Science Foundation through the
%National Centre of Competence in Research Robotics.
We would like to thank all the families who participated to the study.
